Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software

### 2021-08-16 manuscript/manuscript-speed-accuracy.tex
Updating/adding figures, supplementary materials and minor manuscript edits.

\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left

\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{maori}
\graphicspath{{../figures/}}
\usepackage{xr}
\externaldocument{supplementary}

@@

\newcommand{\beginsupplement}{%
        \setcounter{table}{0}
        \renewcommand{\thetable}{S\arabic{table}}%
        \setcounter{figure}{0}
        \renewcommand{\thefigure}{S\arabic{figure}}%
     }

\definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
\definecolor{color2}{RGB}{200,200,200} % Color of the boxes behind the abstract and headings
\definecolor{color3}{RGB}{5,5,5} % Color of the boxes behind the abstract and headings

\JournalInfo{.} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)

\def\numTools{499}
%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
\def\numBenchmarkPubs{69}
%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
\def\numBenchmarks{134}
%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l

\PaperTitle{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}

%1 A meta-analysis of bioinformatics software benchmarks reveals major influences on software accuracy
%2 Analysis of sixty software benchmarks shows the benefits of good software practice 
%3 Assessing the state of the bioinformatic software landscape with a meta-analysis of benchmarks
%4 Sustained software development, not citation metrics, is indicative of accurate bioinformatic software 
%

\Authors{Paul P. Gardner\textsuperscript{1,2}*, James M. Paterson\textsuperscript{3}, Stephanie McGimpsey\textsuperscript{4}, Fatemeh Ashari-Ghomi\textsuperscript{5}, Sinan U. Umu\textsuperscript{6}, Aleksandra Pawlik\textsuperscript{7},
Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
} % Authors

\affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
\affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
\affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
8 Cambridgeshire, CB10 1RQ, UK.}}
\affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
\affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
\affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
\affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation

\affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author

\Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
\newcommand{\keywordname}{Keywords} % Defines the keywords heading name

@@

%\Abstract{In the below we provide additional results for our investigation of computational biology benchmarks.}
\Abstract{Computational biology provides widely used and powerful
  software tools for testing and making inferences about biological
  data. In the face of rapidly increasing volumes of data, heuristic
  methods that trade software speed for accuracy may
  be employed.  We are have studied these trade-offs using the results of a large number of independent software benchmarks, 
  and evaluated whether external factors are indicative of accurate
  software.

We have extracted accuracy and speed ranks from independent benchmarks of
different bioinformatic software tools, and evaluated whether the
speed, author reputation, journal impact, recency and developer efforts are indicative of accuracy.

We found that software speed, author reputation, journal impact, 
number of citations and age are all unreliable predictors of software
accuracy. This is unfortunate because citations, author and journal
reputation are frequently cited reasons for selecting software
tools. However, GitHub-derived records and high version numbers show that 
the accurate bioinformatic software tools
are generally the product of many improvements over time, often from
multiple developers.

We also find
that the field of bioinformatics has a large excess of slow and inaccurate software
tools, and this is consistent across many sub-disciplines. Meanwhile, there are few tools that
are middle-of-road in terms of accuracy and speed trade-offs. We
hypothesise that a form of publication-bias influences the
publication and development of bioinformatic software. In other
words, software that is intermediate in terms of both speed and accuracy may be difficult to publish - possibly due to author,
editor and reviewer practices. This leaves an unfortunate hole in the
literature as the ideal tools may fall into this gap. For example, high accuracy tools are not always useful if years of CPU time are required, 
while high speed is not useful if the results are also inaccurate.  }

\begin{document}

\flushbottom % Makes all text pages the same height
\maketitle % Print the title and abstract box
%\tableofcontents % Print the contents section

\thispagestyle{empty} % Removes page numbering from the first page

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

%\onecolumn
%\beginsupplement

%----------------------------------------------------------------------------------------
%	ARTICLE CONTENTS
%----------------------------------------------------------------------------------------

%%REFERENCES::
%reference Stefan Buchka$^{1}$, Rory Wilson$^2$, Alexander Hapfelmeier$^3$,\\  Paul Gardner$^4$,  Anne-Laure Boulesteix PAPER!!!

%% A comprehensive analysis of the usability and archival stability of omics computational tools and resources
%% https://www.biorxiv.org/content/10.1101/452532v1

%% Tools and best practices for data processing in allelic expression analysis
%% https://genomebiology.biomedcentral.com/articles/10.1186/s13059-015-0762-6

%% Towards reproducible, transparent, and systematic benchmarking of omics computational tools
%% https://osf.io/p8yd9

%% Essential guidelines for computational method benchmarking
%% https://arxiv.org/abs/1812.00661

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
highly-cited software tools include implementations of methods for sequence
alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient, instead usability, ease of installation,
operating system support or other indirect factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
convenient, yet inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

@@

The gold-standard for determining accuracy is for researchers independent
of individual tool development to conduct benchmarking studies, these benchmarks can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021}. Benchmarking typically involves the use a number of
positive and negative control datasets, so that predictions from different software tools can be
partitioned into true or false groups, allowing a variety of metrics to be
used to evaluate performance
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to robustly identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
However, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

@@

With the wide adoption of GitHub ($47\%$ of the
{\color{black}\numTools} tools included in this study could be
linked to a GitHub repository), and consequently quantifiable data on software
development time and intensity indicators, such as the number of contributors to code, number of code
changes and versions is available for these
\cite{ray2014large,Dozmorov:2018,mangul2018comprehensive}.

@@

\begin{figure*}[htb!]
\includegraphics[width=\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  different features of bioinformatic software tools. Spearman’s rho is used to
  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
  version number, GitHub derived activity measures, and 
  the mean
  relative speed and accuracy rankings. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
    different software tool features. 
    The unpermuted correlations are indicated with a red asterix.
    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
\label{fig:allfactors}
\end{figure*}

For each of the publications describing these tools, we have (where
possible) collected the journal's H5-index (\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{Google
Scholar Metrics}), the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that citation metrics
are not static and will change over time. In addition, where possible we also extract the version number, the number
of commits and number of contributors from GitHub repositories.

@@

Data on the number of updates to software tools from GitHub such as
the number of versions, commits and contributors was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
corrected P-values = {\color{black}0.029, 0.042, 0.063}),
Figure~\ref{fig:allfactors}B.  The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A \& Supplementary
Figures~\ref{fig:metricPairs} and~\ref{fig:S2}).  We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).
%The strongest
%association was between accuracy and the relative
%number of citations (Spearman’s rho = {\color{black}0.092,
%P-value = 0.047}). But the effect size is only just over the
%significance threshold and may be the result of multiple-testing.

@@

\begin{figure*}[htb!]
%\includegraphics[scale=0.55]{figure2.pdf}
\centering
\includegraphics[width=0.75\textwidth]{figure2.pdf}
\caption{\textbf{A.} A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Redder
  colours indicate an abundance of software tools in an accuracy and
  speed category, while bluer colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 1,000 random
  permutations of speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 9 classes (a $3\times 3$ grid) that range from
  comparatively slow and inaccurate to comparatively
  fast and accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘$\vartimes$’. \textbf{B.} The z-score distributions from the permutation tests (indicated with the wheat coloured violin plots) compared to 
  the z-score for the observed values for each of the corner and middle square of the heatmap.}
\label{fig:speedaccuracy}
\end{figure*}

@@

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may never-the-less have further uses.
%We also found that slow and inaccurate
%software is generally published earlier than fast and accurate tools
%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
%comparisons with existing software were not required to publish the slow and
%inaccurate tools.

@@

%{\bf AND ETC....}

%----------------------------------------------------------------------------------------
%	REFERENCE LIST
%----------------------------------------------------------------------------------------
\bibliographystyle{unsrt}
\bibliography{references.bib}

\end{document}

>>>>

\documentclass{bmcart}
%\documentclass[fleqn,10pt]{SelfArx} % Document font size and equations flushed left

\usepackage[utf8]{inputenc} %unicode support
\usepackage{float}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[hyphens]{url}
%\usepackage{hyperref}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{todonotes}
\usepackage{maori}
\graphicspath{{../figures/}}
\usepackage{xr}
\externaldocument{supplementary}

@@

%% \definecolor{color1}{RGB}{0,0,90} % Color of the article title and sections
%% \definecolor{color2}{RGB}{200,200,200} % Color of the boxes behind the abstract and headings
%% \definecolor{color3}{RGB}{5,5,5} % Color of the boxes behind the abstract and headings

%\JournalInfo{.} % Journal information ``Journal, Vol. XXI, No. 1, 1-5, 2015''
%\Archive{ } % Additional notes (e.g. copyright, DOI, review/research article)

%%% Put your definitions there:
\startlocaldefs
\def\numTools{499}
%cat meanRankSpeedData.tsv | cut -f 4 | grep -v ^method$ | wc -l
\def\numBenchmarkPubs{69}
%cat speed-vs-accuracy-toolRanks2005-2020.tsv | cut -f 1 | sort  | uniq | grep -E '[0-9]'  | wc -l
\def\numBenchmarks{134}
%cat rawRankSpeedData2005-2020.tsv | cut -f 1 | grep -v ^testId$ | uniq -c | wc -l
\endlocaldefs

%%% Begin BMC format...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%\PaperTitle{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}
\title{Sustained software development, not number of citations or journal choice, is indicative of accurate bioinformatic software}

%% \Authors{Paul P. Gardner\textsuperscript{1,2}*, James M. Paterson\textsuperscript{3}, Stephanie McGimpsey\textsuperscript{4}, Fatemeh Ashari-Ghomi\textsuperscript{5}, Sinan U. Umu\textsuperscript{6}, Aleksandra Pawlik\textsuperscript{7},
%% Alex Gavryushkin\textsuperscript{8}, Michael A Black\textsuperscript{1}
%% } % Authors

%james.paterson@canterbury.ac.nz, srm88@cam.ac.uk, fatgho@food.dtu.dk, sinan.ugur.umu@kreftregisteret.no, PawlikA@landcareresearch.co.nz, alex@biods.org, mik.black@otago.ac.nz

%Authors:
\author[
  addressref={aff1,aff2},                   % id's of addresses, e.g. {aff1,aff2}
  corref={aff1},                       % id of corresponding address, if any
  email={paul.gardner@otago.ac.nz}   % email address
]{\inits{P.P.}\fnm{Paul P.} \snm{Gardner}}
\author[
  addressref={aff3},
  email={james.paterson@canterbury.ac.nz}
]{\inits{J.M.}\fnm{James M.} \snm{Paterson}}
\author[
  addressref={aff4},
  email={srm88@cam.ac.uk}
]{\inits{S.}\fnm{Stephanie} \snm{McGimpsey}}
\author[
  addressref={aff5},
  email={fatgho@food.dtu.dk}
]{\inits{F.}\fnm{Fatemeh} \snm{Ashari-Ghomi}}
\author[
  addressref={aff6},
  email={sinan.ugur.umu@kreftregisteret.no}
]{\inits{S.U.}\fnm{Sinan U.} \snm{Umu}}
\author[
  addressref={aff7},
  email={PawlikA@landcareresearch.co.nz}
]{\inits{A.}\fnm{Aleksandra} \snm{Pawlik}}
\author[
  addressref={aff8},
  email={alex@biods.org}
]{\inits{A.}\fnm{Alex} \snm{Gavryushkin}}
\author[
  addressref={aff1},
  email={mik.black@otago.ac.nz}
]{\inits{M.A.}\fnm{Michael A.} \snm{Black}}

%Affiliations:
%% \affiliation{\textsuperscript{1}\textit{Department of Biochemistry, University of Otago, Dunedin, New Zealand.}} % Author affiliation
%% \affiliation{\textsuperscript{2}\textit{Biomolecular Interaction Centre, University of Canterbury, Christchurch, New Zealand.}}
%% \affiliation{\textsuperscript{3}\textit{Department of Civil and Natural Resources Engineering, University of Canterbury, Christchurch, New Zealand.}}
%% \affiliation{\textsuperscript{4}\textit{Parasites and Microbes, Wellcome Sanger Institute, Wellcome Genome Campus, Hinxton,
%% 8 Cambridgeshire, CB10 1RQ, UK.}}
%% \affiliation{\textsuperscript{5}\textit{Research Group for Genomic Epidemiology, National Food Institute, Technical University of Denmark, Kongens Lyngby, Denmark.}}
%% \affiliation{\textsuperscript{6}\textit{Department of Research, Cancer Registry of Norway, Oslo, Norway.}} % Author affiliation
%% \affiliation{\textsuperscript{7}\textit{New Zealand eScience Infrastructure, 49 Symonds St, Auckland, New Zealand.}}
%% \affiliation{\textsuperscript{8}\textit{Department of Computer Science, University of Otago, Dunedin, New Zealand.}} % Author affiliation

%% \affiliation{*\textbf{Corresponding author}: paul.gardner@otago.ac.nz} % Corresponding author

%% \Keywords{} % Keywords - if you don't want any simply remove all the text between the curly brackets
%% \newcommand{\keywordname}{Keywords} % Defines the keywords heading name

\address[id=aff1]{%                       % unique id
  \orgdiv{Department of Biochemistry,},   % department, if any
  \orgname{University of Otago},          % university, etc
  \city{Dunedin},                         % city
  \cny{New Zealand}                       % country
}
\address[id=aff2]{%
  \orgdiv{Biomolecular Interaction Centre},
  \orgname{University of Canterbury},
  %\street{},
  %\postcode{}
  \city{Christchurch},
  \cny{New Zealand}
}
\address[id=aff3]{%
  \orgdiv{Department of Civil and Natural Resources Engineering},
  \orgname{University of Canterbury},
  \city{Christchurch},
  \cny{New Zealand}
}
\address[id=aff4]{%
  \orgdiv{Parasites and Microbes},
  \orgname{Wellcome Sanger Institute},
  \city{Hinxton},
  \cny{UK}
}
\address[id=aff5]{%
  \orgdiv{Research Group for Genomic Epidemiology},
  \orgname{Technical University of Denmark},
  \city{Kongens Lyngby},
  \cny{Denmark}
}
\address[id=aff6]{%
  \orgdiv{Department of Research},
  \orgname{Cancer Registry of Norway},
  \city{Oslo},
  \cny{Norway}
}
\address[id=aff7]{%
%  \orgdiv{Department of Research},
  \orgname{Manaaki Whenua - Landcare Research},
  \city{Lincoln},
  \cny{New Zealand}
}
\address[id=aff8]{%                       
  \orgdiv{Department of Computer Science},   
  \orgname{University of Otago},          
  \city{Dunedin},                         
  \cny{New Zealand}                       
}

\end{fmbox}% comment this for two column layout

@@

%\Abstract{
\begin{abstractbox}
  \begin{abstract} % abstract
    \parttitle{Background:} Computational biology provides widely used and
    powerful software tools for testing and making inferences about
    biological data. In the face of rapidly increasing volumes of
    data, heuristic methods that trade software speed for accuracy may
    be employed.  We are have studied these trade-offs using the
    results of a large number of independent software benchmarks, and
    evaluated whether external factors are indicative of accurate
    software.

    \parttitle{Method:} We have extracted accuracy and speed ranks from
    independent benchmarks of different bioinformatic software tools,
    and evaluated whether the speed, author reputation, journal
    impact, recency and developer efforts are indicative of accuracy.

    \parttitle{Results:} We found that software speed, author reputation,
    journal impact, number of citations and age are all unreliable
    predictors of software accuracy. This is unfortunate because
    citations, author and journal reputation are frequently cited
    reasons for selecting software tools. However, GitHub-derived
    records and high version numbers show that the accurate
    bioinformatic software tools are generally the product of many
    improvements over time, often from multiple developers.

    We also find that the field of bioinformatics has a
    large excess of slow and inaccurate software tools, and this is
    consistent across many sub-disciplines. Meanwhile, there are few
    tools that are middle-of-road in terms of accuracy and speed
    trade-offs. We hypothesise that a form of publication-bias
    influences the publication and development of bioinformatic
    software. In other words, software that is intermediate in terms
    of both speed and accuracy may be difficult to publish - possibly
    due to author, editor and reviewer practices. This leaves an
    unfortunate hole in the literature as the ideal tools may fall
    into this gap. For example, high accuracy tools are not always
    useful if years of CPU time are required, while high speed is not
    useful if the results are also inaccurate.
%}
\end{abstract}

%% \begin{keyword}
%% \kwd{sample}
%% \kwd{article}
%% \kwd{author}
%% \end{keyword}

\end{abstractbox}
\end{frontmatter}

%\begin{document}
%\flushbottom % Makes all text pages the same height
%\maketitle % Print the title and abstract box
%\tableofcontents % Print the contents section
%\thispagestyle{empty} % Removes page numbering from the first page

\section*{Background}
Computational biology software is widely used and has produced some of
the most cited publications in the entire scientific corpus
\cite{Perez-Iratxeta2007-lv,Van_Noorden2014-kc,Wren2016-xy}. These
highly-cited software tools include implementations of methods for sequence
alignment and homology inference \cite{Altschul1990-ht,Thompson1994-eu,Thompson1997-rl,Altschul1997-ga},
phylogenetic analysis \cite{Felsenstein1985-lj,Saitou1987-zl,Posada1998-qq,Ronquist2003-yh,Tamura2007-ei}, 
biomolecular structure analysis \cite{Sheldrick1990-kc,Sheldrick2008-xy,Jones1991-ik,Laskowski1993-vi,Otwinowski1997-xj}, and
visualization and data collection \cite{Kraulis1991-lt,Berman2000-to}. 
However, the popularity of a
software tool does not necessarily mean that it is accurate or
computationally efficient{\color{red};} instead usability, ease of installation,
operating system support or other indirect factors may play a greater role in a software
tool's popularity. Indeed, there have been several notable incidences where
convenient, yet inaccurate software has caused considerable harm
\cite{leveson1993investigation,cummings2020regulating,Herkert:2020}.
%Fergusson covid predictions?

@@

The gold-standard for determining accuracy is for researchers independent
of individual tool development to conduct benchmarking studies{\color{red};} these benchmarks can serve a useful role in
reducing the over-optimistic reporting of software accuracy
\cite{Boulesteix2010-te,Jelizarow2010-zf,Weber:2019} and the self-assessment trap
\cite{Norel2011-cq,Buchka:2021a}. Benchmarking typically involves the use a number of
positive and negative control datasets, so that predictions from different software tools can be
partitioned into true or false groups, allowing a variety of metrics to be
used to evaluate performance
\cite{Egan1975-nd,Hall2012-kg,Weber:2019}.
%Some benchmarks now use live, or
%frequently updated, results to indicate the latest developments in
%software performance
%\cite{Bujnicki2001-xr,Puton2014-hy,Barton_undated-er}.
The aim of these benchmarks is to robustly identify tools that
make acceptable compromises in terms of balancing speed with
discriminating true and false predictions, and are therefore suited
for wide adoption by the community.

@@

A number of possible predictors of software quality are used by the
community of computational biology software users \cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. 
Some accessible, quantifiable and frequently used proxies for identifying high quality
software include: \textbf{1. Recency:} recently published software
tools may have built upon the results of past work, or be an update to
an existing tool. Therefore these may be more accurate and
faster. \textbf{2. Wide adoption:} a software tool may be widely used
because it is fast and accurate, or because it is well-supported and
user-friendly. In fact,``large user base'', ``word-of-mouth'',
``wide-adoption'', ``personal recommendation'', and ``recommendation from a
close colleague'', were frequent responses to surveys of ``how do
scientists select software?''
\cite{Hannay2009-cf,Joppa2013-vj,Loman2015-bw}. \textbf{3. Journal
impact:} {\color{red}many believe that} high profile journals are run by editors and reviewers who
carefully select and curate the best manuscripts. Therefore, high
impact journals may be more likely to select manuscripts describing
good software \cite{Garfield1955-wf}. \textbf{4. Author/Group
reputation:} the key to any project is the skills of the people
involved, including maintaining a high collective intelligence
\cite{Joppa2013-vj,Woolley2010-ld,Cheruvelil2014-xn}. As a
consequence, an argument could be made that well respected and
high-profile authors may write better software
\cite{Hirsch2005-mt,Bornmann2008-il}. \textbf{5. Speed:} software tools
frequently trade accuracy for speed. For example, heuristic
software such as the popular homology search tool, BLAST, compromises
the mathematical guarantee of optimal solutions for more speed
\cite{Altschul1990-ht,Altschul1997-ga}. Some researchers may naively
interpret this fact as implying that slower software is likely to be more
accurate. But speed may also be influenced by the programming language
\cite{Fourment2008-vl}, and the level of hardware optimisation
\cite{Farrar2007-ky,Dematte2010-ph};
{\color{red}however}, the specific method of implementation generally has a greater impact (e.g., brute-force approaches versus
rapid and sensitive pre-filtering
\cite{Schaeffer1989-mu,Papadimitriou_undated-bo,leiserson2020there}).
%\todo{This paper
%https://science.sciencemag.org/content/368/6495/eaam9744.full
%shows a thousands-fold increase in speed when parallelisation
%is implemented with things like cache, architecture, etc.
%mind VS naive parallelisation as in Python, R, etc.}

@@

With the wide adoption of GitHub ($47\%$ of the
{\color{black}\numTools} tools included in this study could be
linked to a GitHub repository), and consequently quantifiable data on software
development time and intensity indicators, such as the number of contributors to code, number of code
changes and versions is available for these
\cite{ray2014large,Dozmorov:2018,Mangul:2019}.

@@

\begin{figure*}[htb!]
\includegraphics[width=0.9\textwidth]{figure1.pdf}
\caption{\textbf{A.} A heatmap indicating the relationships between
  different features of bioinformatic software tools. Spearman’s rho is used to
  infer correlations between metrics such as citations based metrics, the year and relative age of publication, 
  version number, GitHub derived activity measures, and 
  the mean
  relative speed and accuracy rankings. Red colours
  indicate a positive correlation, blue colours indicate a negative
  correlation. Correlations with an P-value less than 0.05 (corrected for multiple-testing using the Benjami-Hochberg method) are
  indicated with a `$\vartimes$' symbol. The correlations with accuracy are illustrated in
  more detail in \textbf{B}, the relationship between speed and
  accuracy is shown in more detail in \textbf{Figure 2}.
  %The
  %dendrogram was computed using the default ‘heatmap.2 function in R,
  %which computes a Euclidean distance matrix and a complete-linkage
  %hierarchical clustering.
  \textbf{B.} Violin plots of Spearman's correlations between permuted accuracy ranks and 
    different software tool features. 
    The unpermuted correlations are indicated with a red asterix.
    For each benchmark, 1,000 permuted sets of accuracy and speed ranks were generated, and the ranks were normalised to lie between 0 and 1 (see Methods for details).
  Circled asterixs are significant (empirical P-value $< 0.05$, corrected for multiple-testing using the Benjami-Hochberg method).}
\label{fig:allfactors}
\end{figure*}

For each of the publications describing these tools, we have (where
possible) collected the journal's H5-index (
%\href{https://scholar.google.co.nz/citations?view\_op=top\_venues\&hl=en}{
Google Scholar Metrics
%}
), the maximum H-index and
corresponding M-indices \cite{Hirsch2005-mt} for the corresponding
authors for each tool, and the number of times the publication(s)
associated with a tool has been cited using Google Scholar (data
collected over a 6 month period in late 2020). Note that citation metrics
are not static and will change over time. In addition, where possible we also extract the version number, the number
of commits and number of contributors from GitHub repositories.

@@

Data on the number of updates to software tools from GitHub such as
the number of versions, commits and contributors was significantly
correlated with software accuracy (respective Spearman’s rhos =
{\color{black}0.15, 0.22, 0.21,} and respective Benjamini \& Hochberg
corrected P-values = {\color{black}0.029, 0.042, 0.063}),
Figure~\ref{fig:allfactors}B.  The significance of these features was
further confirmed with a permutation test
(Figure~\ref{fig:allfactors}B).  These features were not correlated
with speed however (see Figure~\ref{fig:allfactors}A {\color{red}\& Supplementary
Figures~S5 and~S6}%\ref{fig:S2}
).  We also found that
reputation metrics such as citations, author and journal H-indices,
and the age of tools were generally \textbf{not} correlated with
either tool accuracy or speed (Figure~\ref{fig:allfactors}A\&B).

%The strongest
%association was between accuracy and the relative
%number of citations (Spearman’s rho = {\color{black}0.092,
%P-value = 0.047}). But the effect size is only just over the
%significance threshold and may be the result of multiple-testing.

@@

\begin{figure*}[htb!]
%\includegraphics[scale=0.55]{figure2.pdf}
\centering
\includegraphics[width=0.95\textwidth]{figure2.pdf}
\caption{\textbf{A.} A heatmap indicating the relative paucity or abundance of
  software in the range of possible accuracy and speed rankings. Redder
  colours indicate an abundance of software tools in an accuracy and
  speed category, while bluer colours indicate scarcity of software in
  an accuracy and speed category. The abundance is quantified using a
  Z-score computation for each bin, this is derived from 1,000 random
  permutations of speed and accuracy ranks from each
  benchmark. Mean normalised ranks of accuracy and speed have been
  binned into 9 classes (a $3\times 3$ grid) that range from
  comparatively slow and inaccurate to comparatively
  fast and accurate. Z-scores with a P-value less than 0.05 are indicated
  with a ‘$\vartimes$’. \textbf{B.} The z-score distributions from the permutation tests (indicated with the wheat coloured violin plots) compared to 
  the z-score for the observed values for each of the corner and middle square of the heatmap.}
\label{fig:speedaccuracy}
\end{figure*}

@@

We found the lack of a correlation between software speed and accuracy
surprising.  The slower software tools are overrepresented at both
high and low levels of accuracy (Figure~\ref{fig:speedaccuracy}).
In addition, there is an large under-representation of software that has
intermediate levels of both accuracy and speed. A possible explanation
for this is that bioinformatic software tools are bound by a form of
publication-bias \cite{Boulesteix2015-am,Nissen:2016}. That is, the
probability that a study being published is influenced by the results
it contains \cite{sterling1995publication}. The community of
developers, reviewers and editors may be unwilling to publish software
that is not highly ranked on speed or accuracy. If correct, this may
have unfortunate consequences as these tools may {\color{red}nevertheless} have further uses.
%We also found that slow and inaccurate
%software is generally published earlier than fast and accurate tools
%{\color{black}(P=0.007, one-tailed Wilcoxon test)}, therefore
%comparisons with existing software were not required to publish the slow and
%inaccurate tools.

@@

\section*{Funding}
PPG is supported by a Rutherford Discovery Fellowship,
administered by the Royal Society Te Ap\=arangi, 
PPG, AG and MAB acknowledge support from a Data Science Programmes grant
(UOAX1932).
